Natural transformations

### 2017-07-27 papers/base-3/paper-3.tex

\section*{Equivalence of Categories}
We now want to examine when two categories are the same.
\begin{exmp}
  \[
    \begin{tikzcd}
      \catA: & & \catB:\\
      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
                          & & B' \arrow[bend left]{u} \arrow[loop left]
    \end{tikzcd}
  \]
\end{exmp}
todo: explain example

\begin{defn}
  \label{def:eqv}
  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
  \begin{itemize}
    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
  \end{itemize}
  The functors $F$ and $G$ are called \emph{equivalences}.
  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
\end{defn}

\begin{defn}
  \label{def:surj-on-obj}
  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
\end{defn}

>>>>

\section*{Equivalence of Categories}
We now want to examine when two categories are the same.
To that end let us consider the following example.
\begin{exmp}
  \label{exmp:eqv}
  We take two discrete categories $\catA$ and $\catB$ and define two functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ as depicted below.
  They operate on morphisms in the only possible way.
  \[
    \begin{tikzcd}
      \catA:                                               & & \catB:\\
      A \arrow[loop left] \arrow[dashed, bend left]{rr}{F} & & B \arrow[bend left]{d} \arrow[loop above] \arrow[dashed]{ll}{G}\\
                                                           & & B' \arrow[bend left]{u} \arrow[loop below] \arrow[dashed]{llu}{G}
    \end{tikzcd}
  \]
  The two objects $B$ and $B'$ of $\catB$ are isomorphic.
  By "merging" them we obtain the category $\catA$.
  We now examine how this merging can be described by $F$ and $G$.
  We observe that $G \circ F = 1_{\catA}$.
  But $F \circ G$ is only isomorphic to $1_{\catB}$.
  This isomorphism between the composition of $F$ and $G$ leads to the following definition of equivalence.
\end{exmp}

\begin{defn}
  \label{def:eqv}
  Two categories $\catA$ and $\catB$ are \demph{equivalent} if:
  \begin{itemize}
    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
  \end{itemize}
  The functors $F$ and $G$ are called \demph{equivalences}.
  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
\end{defn}

\begin{defn}
  \label{def:surj-on-obj}
  A functor $F \from \catA \to \catB$ is \demph{essentially surjective on objects}
  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
\end{defn}
